Naturalness in U(1) extensions of the MSSM

### 2014-12-17 paper/e6ssm-tuning-paper.tex
Replaced 1-loop beta for lambda
Moved scan details from section on numerical procedure to results section

Using the approach outlined above, we have scanned the low energy parameter space of the MSSM and E$_6$SSM and calculated the fine tuning in each of the models considered. To do so, we implemented the above expressions for computing the fine tuning in a modified version of the E$_6$SSM spectrum generator that was used in \cite{Athron:2013ipa}. This code implemented two-loop RGEs for all parameters except the soft scalar masses. In order to properly include the fine tuning impact of the $SU(3)$ gaugino soft mass $M_3$, we have extended the original code to make use of the two-loop RGEs generated by SARAH \cite{Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} and FlexibleSUSY \cite{Athron:2014yba}. The CP-even Higgs masses are calculated including the leading one-loop effective potential contributions given in \cite{Athron:2009bs} and the leading two-loop contributions. To scan over the MSSM parameter space, the equivalent MSSM fine tuning expressions were implemented into a modifed version of SOFTSUSY 3.3.10 \cite{Allanach:2001kg,Allanach:2013kza}. For consistency with the results produced in the $E_6$ models, and for computational speed, for our main scans only the dominant one- and two-loop corrections to the CP-even Higgs masses were included. The parameter ranges that were scanned over are presented in Table \ref{tab:scanranges}. With the exception of the stop mixing $A_t$, which was logarithmically scanned, all parameters were linearly scanned. All other parameters were decoupled by setting the remaining soft masses to values of $5$ TeV, while the first and second generation Yukawa couplings and their associated soft trilinears were taken to vanish. The $U(1)$ gaugino soft mass $M_1$ was fixed to $M_1=300$ GeV, and we fix $M_3=2000$ GeV. The input scale $M_X$ was set to $M_X=20$ TeV for both the MSSM and the $E_6$ models. Finally, the fine tuning was evaluated at the scale $Q=M_{\textrm{SUSY}}=\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$, where $m_{\tilde{t}_{1,2}}$ are the running $\overline{\textrm{DR}}$ stop masses evaluated at $Q=M_{\textrm{SUSY}}$. 
\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cc}
MSSM & E$_6$SSM \\
\hline
$2 \leq \tan\beta \leq 50$& $2 \leq \tan\beta \leq 50$ \\
$-1\textrm{ TeV } \leq \mu \leq 1 \textrm{ TeV}$ & $-3 \leq \lambda \leq 3$\\
$-1\textrm{ TeV } \leq B \leq 1\textrm{ TeV}$ & $-10\textrm{ TeV } \leq A_\lambda \leq 10\textrm{ TeV}$ \\
$ 200 \textrm{ GeV } \leq m_{Q_3} \leq 2000 \textrm { GeV}$ & $ 200 \textrm{ GeV } \leq m_{Q_3} \leq 2000 \textrm { GeV}$\\
$ 200 \textrm{ GeV } \leq m_{u_3} \leq 2000 \textrm { GeV}$ & $ 200 \textrm{ GeV } \leq m_{u_3} \leq 2000 \textrm { GeV}$\\
$ -10 \textrm{ TeV } \leq A_t \leq 10 \textrm { TeV}$ & $ -10 \textrm{ TeV } \leq A_t \leq 10 \textrm { TeV}$\\
$M_2=100\textrm{ GeV, } 1050\textrm{ GeV, } 2000 \textrm{ GeV}$ & $M_2=100\textrm{ GeV, } 1050\textrm{ GeV, } 2000 \textrm{ GeV}$ \\
\end{tabular}
\end{ruledtabular}
\caption{The parameters scanned over and the ranges of values used in the MSSM and the $E_6$ models.}
\label{tab:scanranges}
\end{table}
\section{\label{sec:results}Results}
% Our results. Plots to go in here:
%    - tuning as a function of stop mass for the MSSM and E6SSM
Many recent papers have focussed on light stops as allowing natural
SUSY and there has been a lot of theoretical effort to find SUSY
models with a $125$ GeV Higgs bosons while still having light stops
and experimental efforts to search for light stops.  This is entirely
appropriate, however the fine tning problem from the stops depends
crucially on the RG evolution from the GUT scale, since this is how
the squark masses feed into the soft Higgs masses that appear in the
EWSB conditions.

@@

Since the stop mass does not have a large impact on the fine tuning
when the cutoff scale is very low we can use this to see more clearly
the impact of the $Z^\prime$ mass on fine tuning. To do se we cjoose a
cutoff of [insert cutoff] look at the fine tuning for two different
values of the $Z^\prime$ mass, $2.5$ TeV, which is just above the
current limits and $4.5$ TeV which could be in reach of the LHC run
II, and compare these to the MSSM.

For this we also fix [what we fix + explain different choices to
  before and their impact].
\begin{figure}[h]
\begin{center}
\resizebox{!}{10cm}{%
\includegraphics{all_models_tuning_vs_higgs.png}}
\caption{Scatter plot of fine tuning vs lightest Higgs mass for the
  MSSM (green), E$_6$SSM with $M_{Z^\prime} = 2.5$ TeV (blue) and
  E$_6$SSM with $M_{Z^\prime} = 4.5$ TeV (red).}
\label{Fig:e6ssmvsmssm}
\end{center}
\end{figure}
In Fig \ref{Fig:e6ssmvsmssm} we see that as expected the minimum
tuning in the model for the E$_6$SSM is increased by the mass of the
$Z^\prime$ boson.  Since with such a low cutoff the stop masses have
little impact on the tuning, the tuning here also does not depend much
on the Higgs mass.  However the tuning from the $Z^\prime$ mass
appears already at tree level and is unavoidable regardless of the
cutoff scale chosen.

@@

\beta_\lambda^{(1)}&=,\label{eq:USSMLambdaBetaOneLoop}\\
\beta_\lambda^{(2)}&=.\label{eq:USSMLambdaBetaTwoLoop}

@@

>>>>

Using the approach outlined above, we are able to scan the low energy parameter space of the MSSM and E$_6$SSM and calculate the fine tuning in each. To do so, we implemented the above expressions for computing the fine tuning in a modified version of the E$_6$SSM spectrum generator that was used in \cite{Athron:2013ipa}. This code implemented two-loop RGEs for all parameters except the soft scalar masses. In order to properly include the fine tuning impact of the $SU(3)$ gaugino soft mass $M_3$, we have extended the original code to make use of the two-loop RGEs generated by SARAH \cite{Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} and FlexibleSUSY \cite{Athron:2014yba}. The CP-even Higgs masses are calculated including the leading one-loop effective potential contributions given in \cite{Athron:2009bs} and the leading two-loop contributions. To scan over the MSSM parameter space, the equivalent MSSM fine tuning expressions were implemented into a modifed version of SOFTSUSY 3.3.10 \cite{Allanach:2001kg,Allanach:2013kza}. For consistency with the results produced in the $E_6$ models, and for computational speed, for our main scans only the dominant one- and two-loop corrections to the CP-even Higgs masses were included. Finally, in all of the results below the fine tuning was evaluated at the scale $Q=M_{\textrm{SUSY}}=\sqrt{m_{\tilde{t}_1}m_{\tilde{t}_2}}$, where $m_{\tilde{t}_{1,2}}$ are the running $\overline{\textrm{DR}}$ stop masses evaluated at $Q=M_{\textrm{SUSY}}$.

\section{\label{sec:results}Results}
% Our results. Plots to go in here:
%    - tuning as a function of stop mass for the MSSM and E6SSM
Many recent papers have focussed on light stops as allowing natural
SUSY and there has been a lot of theoretical effort to find SUSY
models with a $125$ GeV Higgs boson while still having light stops
and experimental efforts to search for light stops.  This is entirely
appropriate, however the fine tning problem from the stops depends
crucially on the RG evolution from the GUT scale, since this is how
the squark masses feed into the soft Higgs masses that appear in the
EWSB conditions.

@@

Since the stop mass does not have a large impact on the fine tuning when the 
cutoff scale is very low we can use this to see more clearly the impact of the
$Z^\prime$ mass on fine tuning. To do so we select a fixed low cutoff of $M_X=20$ TeV and compare the
fine tuning for two different values of the $Z^\prime$ mass. We choose to look at
$M_{Z^\prime}=2.5$ TeV, which is just above the current limits, and $M_{Z^\prime}=4.5$ TeV, 
which should be in reach in run II at the LHC. We also compare the fine tuning calculated
in each case to the tuning in the MSSM. For this, we have performed a parameter space scan in both the 
MSSM and E$_6$SSM, scanning over those parameters that appear at tree level in the EWSB conditions in the
models. While the contribution from large stop masses to the fine tuning is small for such a low cut-off scale, the 
stop contributions to the effective potential can play a significant role in reducing the fine tuning.
For this reason it is still important to properly treat the tuning associated with stop contributions to the one-loop effective potential, and so we also scan over the soft masses $m_{Q_3}^2$, $m_{u_3}^2$ and the stop mixing $A_t$.
The relevant parameters and ranges that were scanned over are summarised in Table \ref{tab:scanranges}. With the exception of the stop mixing $A_t$ and the soft trilinear $A_\lambda$, which are logarithmically scanned, all parameters are linearly scanned. In both the MSSM and E$_6$SSM, we then decouple all other parameters by setting the remaining soft masses to values of $5$ TeV, while the first and second generation Yukawa couplings and their associated soft trilinears are taken to vanish. The $U(1)$ gaugino soft mass $M_1$ was fixed to $M_1=300$ GeV, and we fix $M_3=2000$ GeV. Additionally, in the E$_6$SSM the $U(1)_N$ gaugino soft mass $M_1'$ is held fixed with $M_1'=M_1=300$ GeV, and $\mu^\prime=5$ TeV.
\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cc}
MSSM & E$_6$SSM \\
\hline
$2 \leq \tan\beta \leq 50$& $2 \leq \tan\beta \leq 50$ \\
$-1\textrm{ TeV } \leq \mu \leq 1 \textrm{ TeV}$ & $-3 \leq \lambda \leq 3$\\
$-1\textrm{ TeV } \leq B \leq 1\textrm{ TeV}$ & $-10\textrm{ TeV } \leq A_\lambda \leq 10\textrm{ TeV}$ \\
$ 200 \textrm{ GeV } \leq m_{Q_3} \leq 2000 \textrm { GeV}$ & $ 200 \textrm{ GeV } \leq m_{Q_3} \leq 2000 \textrm { GeV}$\\
$ 200 \textrm{ GeV } \leq m_{u_3} \leq 2000 \textrm { GeV}$ & $ 200 \textrm{ GeV } \leq m_{u_3} \leq 2000 \textrm { GeV}$\\
$ -10 \textrm{ TeV } \leq A_t \leq 10 \textrm { TeV}$ & $ -10 \textrm{ TeV } \leq A_t \leq 10 \textrm { TeV}$\\
$M_2=100\textrm{ GeV, } 1050\textrm{ GeV, } 2000 \textrm{ GeV}$ & $M_2=100\textrm{ GeV, } 1050\textrm{ GeV, } 2000 \textrm{ GeV}$ \\
\end{tabular}
\end{ruledtabular}
\caption{The parameters scanned over and the ranges of values used in the MSSM and the $E_6$ models.}
\label{tab:scanranges}
\end{table}

%   Since the stop mass does not have a large impact on the fine tuning
%   when the cutoff scale is very low we can use this to see more clearly
%   the impact of the $Z^\prime$ mass on fine tuning. To do se we choose a
%   cutoff of  look at the fine tuning for two different
%   values of the $Z^\prime$ mass, $2.5$ TeV, which is just above the
%   current limits and $4.5$ TeV which could be in reach of the LHC run
%   II, and compare these to the MSSM.

%   For this we also fix [what we fix + explain different choices to
%   before and their impact].
\begin{figure}[h]
\begin{center}
\resizebox{!}{10cm}{%
\includegraphics{all_models_tuning_vs_higgs.png}}
\caption{Scatter plot of fine tuning vs lightest Higgs mass for the
  MSSM (green), E$_6$SSM with $M_{Z^\prime} = 2.5$ TeV (blue) and
  E$_6$SSM with $M_{Z^\prime} = 4.5$ TeV (red).}
\label{Fig:e6ssmvsmssm}
\end{center}
\end{figure}
In Fig \ref{Fig:e6ssmvsmssm} we see that as expected the minimum
tuning in the model for the E$_6$SSM is increased by the mass of the
$Z^\prime$ boson.  Since with such a low cutoff the stop masses have
little impact on the tuning, the tuning here also does not depend much
on the Higgs mass.  However the tuning from the $Z^\prime$ mass
appears already at tree level and is unavoidable regardless of the
cutoff scale chosen.

@@

\beta_\lambda^{(1)}&=\lambda  \left[2\lambda ^2+2 \Sigma_\lambda +3 \Sigma_\kappa +3 y_t^2+3 y_b^2+y_{\tau }^2-3 g_2^2-\frac{3}{5} g_1^2 -2 \left ( \tilde{Q}_1^2 + \tilde{Q}_2^2 + \tilde{Q}_S^2\right )g_1'^2 \right],\label{eq:USSMLambdaBetaOneLoop}\\
\beta_\lambda^{(2)}&=\lambda \left [ \left ( \right )g_1'^4\right ],\label{eq:USSMLambdaBetaTwoLoop}

@@

\Sigma_\lambda=\lambda_1^2+\lambda_2^2+\lambda_3^2, \qquad \Sigma_\kappa=\kappa_1^2+\kappa_2^2+\kappa_3^2.
